---
title: "R Notebook"
output: html_notebook
--- 

## Loading Dataset

```{r}
df = read.csv("C:/Users/RIDZZ ZOLDYCK/Desktop/SY SEM2/DS/educ_exp_ha.csv")
```


```{r}
head(df,5)
```

```{r}
colnames(df)

```

```{r}
str(df)
```

## Removing insignificant columns 

```{r taking required columns}
data = df[ , c(2,3,4,5)]
colnames(data)
```
```{r dplyr package}
library(dplyr)
```

```{r renaming X as Country}
data = rename(data, Country = X)  # dplyr
colnames(data)
```

```{r}
data
```

## Extracting Public Expenditure Data

```{r}
new_data = data[data['Series'] == "Public expenditure on education (% of GDP)" , ] 
new_data
```



```{r}

summary(new_data) #checking for NA values

```





```{r}

#unique(new_data$Country)
# 
# country = unique(new_data$Country)
# tmp = c()
# k = 1
# #class(i)
# 
# 
# for (i in country){
#   
#   tmp1 = length(new_data[new_data$Country == i , 1 ])
#   
#   tmp[k] = tmp1 
#   
#   k = k + 1
#   
# }
# 
# #tmp
# 
# coun_data = cbind(country , tmp)
# #coun_data
# 
# # dataframe containing countries and their number of instances is created
# class(coun_data)
# coun_data = as.data.frame(coun_data)
# 
# 
# arrange(coun_data , desc(tmp))
# 

```




```{r}

# sel_coun = c("Burundi" , "El Salvador" , "Estonia" , "Jamaica" , "Madagascar" , "Mongolia" )
#   
# for (i in sel_coun){
#   
#   yr = new_data$Year[new_data$Country == i]
#   print(i)
#   print(yr)
# }

```





### Our data has 190+ countries
### Since all these are categorical variables it isnt practical to include 190+ categorical variables
### So we will be selecting 10 countries and do our processing and analysis on them


```{r}
# Countries chosen - India , United Kingdom , Canada , Russian Federation , Republic of Korea (South Korea) , United States of America , Israel ,  Japan , Spain , Germany

coun = c( 'India' , 'United Kingdom' , 'Canada' , 'Russian Federation' , 'Republic of Korea' , 'United States of America' , 'Israel' ,  'Japan' , 'Spain' , 'Germany')

ex_data_ = NULL
flag = 0

for (i in coun)
{
    if(flag == 0){
      
      ex_data_ = new_data[new_data['Country'] == i ,]
      flag = flag + 1 
    }else{
      
      ex_data_ = rbind(ex_data_ , new_data[new_data['Country'] == i ,] )
    }
    
}


# ex_data = new_data[new_data['Country'] == "India" ,]
# ex_data

unique(ex_data_$Country)

```

```{r}
# contains data of the selected 10 countries
ex_data_


```


```{r}
# Now that we have the data of the required countries we will remove the Series column as all of them have the same value [Public expenditure on education (% of GDP)]

# Also we will rename the Value for better understanding

ex_data_ = subset(ex_data_, select = -c(3) )

```


```{r}
ex_data_
```

## Data Visualization 


```{r}

# Since all countries have different values in Year , we cannot have a common x axis for a line plot
# Hence we ll make a different like plot for each country 

library(ggplot2)

#scale_colour_manual()



# data_india = ex_data_[ex_data_$Country == "India" , ]
# #data_india
# 
# 
# india_line_plot = ggplot(data_india , aes(Year , Value , group = 1) ) + geom_point(color = "red" , size = 5) + geom_line(linetype = "dashed" , color = "blue" , size = 1.5)
# 
# 
# india_line_plot



p = ggplot(ex_data_ , aes(x = Year , y = Value) ) + geom_point(aes( color = factor(Country)  ) , size = 2.5 , show.legend = TRUE) + scale_colour_manual("Country" , values = c("blue4", "red3", "gold3" , "darkorchid4" , "darkorange3" , "chartreuse3" , "steelblue1" , "indianred1" , "grey3" , "aquamarine")) + labs(x = "Year" , y = "% of GDP in Education" , title = "Country wise GDP Expenditure on Education")
p = p + coord_cartesian(xlim = c(2000, 2017), ylim = c(3, 6)) +  theme(plot.title = element_text(hjust = 0.5) , panel.background = element_rect(fill = "khaki3"))

p


# canada , india , russian federation , united kingdom , canada , spain , israel 

```


# We will now plot individual countries to have a better country wise insight 


```{r}

# India 


data_india = ex_data_[ex_data_$Country == "India" , ]
#data_india


india_line_plot = ggplot(data_india , aes(Year , Value ) ) + geom_point(color = "blue" , size = 3) + geom_line(linetype = "dashed" , color = "red" , size = 1)

india_line_plot = india_line_plot + labs(x = "Year" , y = "% of GDP in Education" , title = "India's GDP Expenditure on Education") + theme(plot.title = element_text(hjust = 0.5) , panel.background = element_rect(fill = "lightskyblue1"))


india_line_plot




```


```{r}

# Russian Federation 

data_russia = ex_data_[ex_data_$Country == "Russian Federation" , ]
data_russia


russia_line_plot = ggplot(data_russia , aes(Year , Value ) ) + geom_point(color = "darkred" , size = 3) + geom_line( linetype = "dotted" , color = "darkslateblue" , size = 1)

russia_line_plot = russia_line_plot + labs(x = "Year" , y = "% of GDP in Education" , title = "Russia's GDP Expenditure on Education") + theme(plot.title = element_text(hjust = 0.5) , panel.background = element_rect(fill = "lightskyblue1"))

russia_line_plot
```


```{r}

# Israel 

data_israel = ex_data_[ex_data_$Country == "Israel" , ]
data_israel

temp_is = data_israel$Year

data_israel$Year = as.factor(data_israel$Year)  # prevents spacing between different bars as it treats them like numbers


israel_bar_plot = ggplot(data_israel , aes(Year ,Value) ) + geom_bar(stat = "identity" , fill = c(
        "deepskyblue4",
        "deeppink3",
        "darkgoldenrod2",
        "darkorchid" ) , width = 0.5) + coord_flip() + geom_text(aes( label = Value ) , hjust = 1.05) + labs(x = "Year" , y = "% of GDP in Education" , title = "Israel's GDP Expenditure on Education") + theme(plot.title = element_text(hjust = 0.5) , panel.background = element_rect(fill = "lightskyblue1"))
    

        
          
        
        
        
        
        
        
israel_bar_plot

data_israel$Year = temp_is
```


```{r}

# United Kingdom

data_uk = ex_data_[ex_data_$Country == "United Kingdom" , ]
data_uk

temp_uk = data_uk$Year

data_uk$Year = as.factor(data_uk$Year)

uk_bar_plot = ggplot(data_uk , aes(Year ,Value) ) + geom_bar(stat = "identity" , fill = c(
        "slateblue2",
        "darksalmon",
        "lightgoldenrod3",
        "palegreen3" ) , width = 0.5) +  geom_text(aes( label = Value ) , vjust = -0.5) + labs(x = "Year" , y = "% of GDP in Education" , title = "United Kingdom's GDP Expenditure on Education") + theme(plot.title = element_text(hjust = 0.5) , panel.background = element_rect(fill = "lightskyblue"))

uk_bar_plot

data_uk$Year = temp_uk

```


```{r}

data_canada = ex_data_[ex_data_$Country == "Canada" , ]
data_canada

```




## Plotting per year GDP Expenditure on Education increase of the countries 

## Formula used - 
    
                    ( max_year_Value - min_year_Value ) / ( max_Year - min_Year )                 

  
```{r}

# India
max_year_india = max(data_india$Year)
#max_year_india
min_year_india = min(data_india$Year)
#min_year_india
max_gdp_india = data_india$Value[data_india$Year == max_year_india]
#max_gdp_india
min_gdp_india = data_india$Value[data_india$Year == min_year_india]
#min_gdp_india

gdp_exp_inc_india = ( max_gdp_india - min_gdp_india ) / ( max_year_india - min_year_india )
gdp_exp_inc_india

# Russia 
max_year_russia = max(data_russia$Year)
#max_year_russia
min_year_russia = min(data_russia$Year)
#min_year_russia
max_gdp_russia = data_russia$Value[data_russia$Year == max_year_russia]
#max_gdp_russia
min_gdp_russia = data_russia$Value[data_russia$Year == min_year_russia]
#min_gdp_russia
gdp_exp_inc_russia = ( max_gdp_russia - min_gdp_russia ) / ( max_year_russia - min_year_russia )
gdp_exp_inc_russia


# Israel
max_year_israel = max(data_israel$Year)
#max_year_israel
min_year_israel = min(data_israel$Year)
#min_year_israel
max_gdp_israel = data_israel$Value[data_israel$Year == max_year_israel]
#max_gdp_israel
min_gdp_israel = data_israel$Value[data_israel$Year == min_year_israel]
#min_gdp_israel

gdp_exp_inc_israel = ( max_gdp_israel - min_gdp_israel ) / ( max_year_israel - min_year_israel )
gdp_exp_inc_israel


# United Kingdom
max_year_uk = max(data_uk$Year)
min_year_uk = min(data_uk$Year)
max_gdp_uk = data_uk$Value[data_uk$Year == max_year_uk]
min_gdp_uk = data_uk$Value[data_uk$Year == min_year_uk]

gdp_exp_inc_uk = ( max_gdp_uk - min_gdp_uk ) / ( max_year_uk - min_year_uk )
gdp_exp_inc_uk


# Canada 
max_year_canada = max(data_canada$Year)
min_year_canada = min(data_canada$Year)
max_gdp_canada = data_canada$Value[data_canada$Year == max_year_canada]
min_gdp_canada = data_canada$Value[data_canada$Year == min_year_canada]

gdp_exp_inc_canada = ( max_gdp_canada - min_gdp_canada ) / ( max_year_canada - min_year_canada )
gdp_exp_inc_canada



```

```{r}

country = c("India" , "Russia" , "Israel" , "UK" , "Canada")
educ_exp_inc = c(gdp_exp_inc_india , gdp_exp_inc_russia , gdp_exp_inc_israel , gdp_exp_inc_uk , gdp_exp_inc_canada)

plot_data = cbind(country , educ_exp_inc)


plot_data = as.data.frame(plot_data)


plot_data$educ_exp_inc = as.numeric(plot_data$educ_exp_inc)
plot_data
```


```{r}


colour = c("deepskyblue4",
        "deeppink3",
        "darkgoldenrod2",
        "darkorchid" ,
        "chartreuse3")

exp_inc_plot = ggplot(plot_data , aes(country , educ_exp_inc) ) + geom_point(size = 4 , color = colour) + geom_segment( aes(x = country , xend = country , y = 0 , yend = educ_exp_inc)  , size = 0.75 , color = colour) 

exp_inc_plot = exp_inc_plot + coord_flip() + labs(x = "Country" , y = "% of GDP" , title = "Increase in Education Expenditure Per Year") + theme(plot.title = element_text(hjust = 0.5) , panel.background = element_rect(fill = "lightskyblue1"))

exp_inc_plot


```


```{r}

reg_data = ex_data_

# from the first plot we know that Republic of Korea , Japan and United States of Amaerica have only 1 row each. Hence we will be removing them 


reg_data = reg_data[  reg_data$Country != 'Republic of Korea'  ,  ]
reg_data = reg_data[  reg_data$Country != 'United States of America'  ,  ]
reg_data = reg_data[  reg_data$Country != 'Japan'  ,  ]

(reg_data$Country)





```


```{r}

# Label Encoding 
library(superml)

lbl = LabelEncoder$new()
reg_data$Country = lbl$fit_transform(reg_data$Country)




```

```{r}
reg_data$Country
#class(reg_data$Country)

```


```{r}
mlr_mdl = lm(Value~Country+Year , reg_data)
summary(mlr_mdl)
```

